---
title: "Untitled"
author: "Ryan Morgan"
date: "December 7, 2016"
output: word_document
---

Make a function to compare someones picks to the key:

```{r}
score <- function(key, picks){
  total = sum(key == picks)
  committee = sum(key[c(35:41)] == picks[c(35:41)])
  semisChamp = sum(key[c(39:41)]== picks[c(39:41)])
  champ = sum(key[41]==picks[41])
  
  score = total + committee * .1 + semisChamp * .01 + champ*.001
  
  return(score)
  
}

score(key2017, picks2017$Ryan)

```

Make a function to make a list of peoples scores:

```{r}
library(tidyverse)

compare <- function(key, picks){
  participants <- length(c(picks[1,]))
  
  
  rankingsList <- data.frame(Participant=colnames(picks), picksCorrect=NA)
  
  
  counter <- 1
  
  while(counter<participants+1){
    
    rankingsList[counter,2] = score(key, picks[,counter])
    counter = counter+1
  }
  
rankingsListNew <- rankingsList %>% arrange(-picksCorrect)
  
 return(rankingsListNew)
}

compare(key2017, picks2017)

```


Make a function to return who the current winner would be


```{r}
winning<- function(key,picks){
 compareStuff <- compare(key,picks)
 
 winner <- compareStuff[1,1]
 
 return(winner)
}

```



Find someone's "Best Case Scenario" for the remaining games

```{r}
bestcase <- function(key, persons_picks){
  bestcase <- key
  
  index_of_zeroes <- which(key==0)
  
  bestcase[c(index_of_zeroes)] = persons_picks[c(index_of_zeroes)]
  
  return(bestcase)
}


```

Use the best case scenario to see if the person wins their own best case:

```{r}
eliminated <- function(key, picks, person){
  
  bestCase <- bestcase(key,picks[,person])
  
  winner <- winning(bestCase,picks)
  
  eliminated <- winner!=colnames(picks)[person]
  return(eliminated)
}

```

Return list of people who are eliminated, based on them not even winning their best case.
```{r}
losers <- function(key,picks){
  participants <- length(picks[1,])
  
  counter = 1
  
  loserLogic <- c(NA)
  
  while(counter < participants+1){
    loserLogic[counter] = eliminated(key,picks,counter)
    counter=counter+1
  }
  
  losers <- colnames(picks)[loserLogic]
return(losers)  
}
```


Make a function to compare how many different are left between two people.

```{r}
differentleft <- function(person1, person2, key){
  futureGames <- key == 0
  person1Future = person1[futureGames]
  person2Future = person2[futureGames]
  
  differentRemaining = sum(person1Future != person2Future)
  
  return(differentRemaining)
  
}

```

Total Different Remaining For all:
```{r}
differentleftAll <- function(picks, key){
  participants <- length(picks[1,])
  
  returnMatrix <- matrix(data = NA, ncol = participants, nrow = participants)
  colnames(returnMatrix) = colnames(picks)
  rownames(returnMatrix) = colnames(picks)
  
  counter1 <- 1
  
  while(counter1 < participants+1){
    counter2 = 1
    while(counter2<participants+1){
      returnMatrix[counter1,counter2]=differentleft(picks[,counter1],picks[,counter2],key)
      counter2=counter2+1
    }
    counter1=counter1+1
  }
  
  returnMatrix <- as.data.frame(returnMatrix)
  
  return(returnMatrix)
}


```

Total Different Function:

```{r}

totalDifferent <- function(person1, person2){
  totalDifferent <- sum(person1 != person2)
  
  return(totalDifferent)
  
}

```


totalDifferent Fucntion for everybody:

```{r}
totalDifferentAll <- function(picks){
    participants <- length(picks[1,])
  
  returnMatrix <- matrix(data = NA, ncol = participants, nrow = participants)
  colnames(returnMatrix) = colnames(picks)
  rownames(returnMatrix) = colnames(picks)
  
  counter1 <- 1
  
  while(counter1 < participants+1){
    counter2 = 1
    while(counter2<participants+1){
      returnMatrix[counter1,counter2]=totalDifferent(picks[,counter1],picks[,counter2])
      counter2=counter2+1
    }
    counter1=counter1+1
  }
  
  returnMatrix <- as.data.frame(returnMatrix)
  
  return(returnMatrix)
}

```

Function For how many games Completed:
```{r}
numColmpleted <- function(key){
  
  gamesPlayed <- sum(key != 0)
  
  return(gamesPlayed)
}

```

Function for how many total Games:

```{r}

numGames <- function(key){
  numGames <- length(key)
  return(numGames)
}
```

Function for how many games remaining:
```{r}
numRemaining <- function (key){
  totalGames <- numGames(key)
  GamesCompleted<- numColmpleted(key)
  
  gamesRemaining <- totalGames - GamesCompleted
  
  return(gamesRemaining)
}

```


Function to put it all together:

```{r}
bowlGameScenarios <- function(picks,key){
  cat("Total Number of Games: ",numGames(key))
  
  cat("\n Games Played: ", numColmpleted(key))
  
  cat("\n Games Remaining: ",numRemaining(key), "\n \n \n")
  
  print (compare(key,picks))
  
  cat("\n Eliminated: ", losers(key , picks))
  
  cat(" \n \n Total Different Matrix: \n")
  
  print(totalDifferentAll(picks=picks))
  
  cat("\n \n \n Different Remaining Matrix: \n")
  
  print(differentleftAll(picks=picks, key=key))
        
  
}
```


The 2017 Situation:

```{r}
Bowls2017 <- read.csv(file="C:\\Users\\Ryan\\Desktop\\Bowl_Games_2017\\2017Bowl.csv",header=TRUE)

key2017 <- Bowls2017$Numeric.Key

key2017[is.na(key2017)] <- 0

picks2017 <- Bowls2017[,c(14:21)]

bowlGameScenarios(picks2017,key2017)

combination_percent_chance_of_winning(picks=picks2017, key=key2017, probs=fpi_probs$Team2_Win_Prob, num_sims=5000, number_of_runs = 2 )

```





Make a function to make all possible combinations of next winners, see who the winner is in those scenarios. 


issue:
in the semifinal, we have 2 vs 3 and 1 vs 4. So scenarios where 2 wins cant have 3 winning the championship. 



Possible semifinal possibilities:

(Oklahoma v Georgia, Clemson v Alabama, national championship)

OC  O
OC  C

OA  O
OA  A

GC  G
GC  C

GA  G
GA  A


(1,1,1)
(1,1,2)
(1,2,2)
(1,2,4)

(2,1,3)
(2,1,1)

(2,2,3)
(2,2,4)

```{r}
all_possible_scenarios <- function(num_games){
n<- num_games

l<- c(rep(list(1:2),n-3))

all_poss<- expand.grid(l)

all_poss <- t(all_poss)

num_scenarios <- 2^(n-3)

semis_1 <- c(1,1,1)
semis_2 <- c(1,1,2)

semis_3<- c(1,2,2)
semis_4 <- c(1,2,4)

semis_5<- c(2,1,3)
semis_6 <- c(2,1,1)

semis_7<-c(2,2,3)
semis_8<-c(2,2,4)

TEST1<- bind_rows(all_poss, semis_1[1],semis_1[2],semis_1[3])

TEST2<- bind_rows(all_poss, semis_2[1],semis_2[2],semis_2[3])

TEST3<- bind_rows(all_poss, semis_3[1],semis_3[2],semis_3[3])

TEST4<- bind_rows(all_poss, semis_4[1],semis_4[2],semis_4[3])

TEST5<- bind_rows(all_poss, semis_5[1],semis_5[2],semis_5[3])

TEST6<- bind_rows(all_poss, semis_6[1],semis_6[2],semis_6[3])

TEST7<- bind_rows(all_poss, semis_7[1],semis_7[2],semis_7[3])

TEST8<- bind_rows(all_poss, semis_8[1],semis_8[2],semis_8[3])

all_poss <- bind_cols(TEST1,TEST2,TEST3,TEST4,TEST5,TEST6,TEST7,TEST8)
return(all_poss)

}

all_possibilities<- all_possible_scenarios(20)
```

Use the FPI to find a % chance that somebody wins. 

Make some simulations:

```{r}
simulate_key <- function(picks, key, probs){

  num_completed = numColmpleted(key = key)
  
  num_games = numGames(key = key)
  
  num_simulations <- num_games - num_completed
  
  sim_probs <- probs[num_completed+1:num_games-3]
  
  simulations <- rbinom(n=num_games-3-num_completed, 1, prob=sim_probs)
  
  simulations = simulations+1
  
  completed_games <- key2017[1:numColmpleted(key2017)]
  
  semis_1 <- c(1,1,1)
  semis_2 <- c(1,1,2)

  semis_3<- c(1,2,2)
  semis_4 <- c(1,2,4)

  semis_5<- c(2,1,3)
  semis_6 <- c(2,1,1)

  semis_7<-c(2,2,3)
  semis_8<-c(2,2,4)
  
  sim_df <- data.frame(S1 = c(completed_games, simulations, semis_1), S2 = c(completed_games, simulations, semis_2), S3 = c(completed_games, simulations, semis_3), S4=c(completed_games, simulations, semis_4), S5 = c(completed_games, simulations, semis_5), S6 = c(completed_games, simulations, semis_6), S7 = c(completed_games, simulations, semis_7), S8 = c(completed_games, simulations, semis_8))
  
  
  return(sim_df)

}


fpi_probs <- read.csv(file="C:\\Users\\Ryan\\Desktop\\Bowl_Games_2017\\fpi_probs.csv",header=TRUE)


simulate_key(picks= picks2017, key=key2017, probs = fpi_probs$Team2_Win_Prob)

```



Now, make a function to do a ton of simulations. 


```{r}
mult_simulations <- function(picks=picks2017, key=key2017, probs=fpi_probs$Team2_Win_Prob, num_sims=10){
  
  ret_df = simulate_key(picks, key, probs)
  
  counter=2
  
  
  while(counter< num_sims+1){
    
    
    new_df = simulate_key(picks, key, probs)
    
    ret_df = bind_cols(ret_df, new_df)
    
    counter=counter+1
  }
  
  return(ret_df)
}

```



Make a function to find the winner for multiple scenarios. 


```{r}
mult_winners <- function(picks, sims){
  num_sims = dim(sims)[2]
  
  ret_list = rep(NA,num_sims)
  
  
  counter=1
  
  while(counter<num_sims+1){
    ret_list[counter] = paste(winning(key=sims[,counter], picks=picks))
    
    counter=counter+1
    
  }
  
  return(ret_list)
  
}



```

A function to put it all together, and just give a percent chance of winning:


```{r}
percent_chance_of_winning <- function(picks, key, probs, num_sims){
  simulations = mult_simulations(picks=picks, key=key, probs=probs, num_sims=num_sims)
  
  
  winners= mult_winners(picks=picks, sims = simulations)
  
  ret_table = table(winners)/(8*num_sims)*100
  
  return(ret_table)
}


```

A function to find combinations of percent chances of winning:

```{r}
combination_percent_chance_of_winning <- function(picks, key, probs, num_sims, number_of_runs){
  
  return_table = percent_chance_of_winning(picks, key, probs, num_sims)
  
  counter=2
  
  while(counter< number_of_runs+1){
    
    return_table = return_table + percent_chance_of_winning(picks, key, probs, num_sims)
    
   counter=counter+1 
  }
  
  
 return_table= return_table/number_of_runs
  
  return(round(return_table,2))
  
}


```